### 2025-05-25 conteudo/methods/methods.tex
add some work 18/05
add some work from 24-05

In the Free Software Development course at University of São Paulo, ministred by the advisor 
of this work at the first semester of 2025, we have proposed to the students the possibility
of mitigating duplications in the IIO subsystem and the AMD display driver found with the Arkanjo tool.
The course had a total of X students, and of that, X students being undergraduate students
and Y begin graduate students.

@@

From the students enrolled in the in the course, X opted for the first alternative and 1 opted for the second
alternative, forming Y groups working on the first alternative and 1 group working on the second alternative. 
These number represent X\% of the students choosing for approching duplications them the other contribution 
alternatives, including more simpler and easy to be done.  It is worth highlighting that none of the students
that choose to mitigate duplications had made a contribution to the Linux kernel before,
meaning that the patch to mitigate the duplication being the first tentative to contribute to the kernel as
newcomers.

The groups have been request to document the experience and approaches to refactor the duplications and the
experience of submiting a patch to the Linux kernel in blogs. We had analyzed these blogs to undertand 
patterns of refactor used for mitigating the duplications, the experience on send the patches and the 
opinions of the subsystem maintainers on the mitigations. Appendix X list the blogs and patches made by the 
student.

For the teaching assistant and the student that executed the Arkanjo tool to explore duplications, 
we have asked them to answer some questions related to the experience of using the tool and what could be
improved. The question asked can be found below and the answers can be found in Appendix Y.

>>>>

In the Free Software Development course at University of São Paulo, ministred by the advisor 
of this work at the first semester of 2025, we have proposed to the students the possibility
of mitigating duplications in the IIO subsystem and the AMD display driver found with the Arkanjo tool.
The course had a total of 26 students, and of that, X students being undergraduate students
and Y begin graduate students.

@@

From the students enrolled in the in the course, 25 students opted for the first alternative and 1 student
opted for the second alternative, forming 11 groups working on the first alternative and 1 group 
working on the second alternative. It is worth highlighting that none of the students
that choose to mitigate duplications had made a contribution to the Linux kernel before,
meaning that the patch to mitigate the duplication being the first tentative to contribute to the kernel as
newcomers.

The groups have been request to document the experience and approaches to refactor the duplications and the
experience of submiting a patch to the Linux kernel in blogs. We had analyzed these blogs to undertand 
patterns of refactor used for mitigating the duplications, the experience on send the patches and the 
opinions of the subsystem maintainers on the mitigations.

For the teaching assistant and the student that executed the Arkanjo tool to explore duplications, 
we have asked them to answer some questions related to the experience of using the tool and what could be
improved. The questions asked is the following:

### 2025-05-25 conteudo/results/participant.tex
add some work from 24-05

The initial plan of this research was to approach and sent a large number of mitigations to
collect a broad number of artifacts from the driver community, but the slow process to send a patch
makes this plan unviable. Thus, we opted to redirect this research to work with students and refactoring duplicated function
pairs found by the tool directly in the IIO and the AMD Display driver, without passing through the whole process 
of the systematic approach proposed.
This alternative allowed us to collect artifacts from the students experiences and findings, and not be 
to dependable of the iteraction with the kernel community.
A additional advantage is to accelerate the iteraction with the kernel community
because we are not applying the whole systematic, meaning that the code changes became smaller.

\subsection{OLD}

In the first function pair, completing the refactoring requires the expertise of a maintainer for the AMD Display driver. 
This observation is because duplicated code artifacts involve copy-pasted code that imports different configuration files.
The design of the AMD Display driver configuration files necessitates importing specific configurations at compile time, 
which prevents a generic approach. Consequently, refactoring this pair requires changes in how the AMD Display driver 
manages configuration files -- a complex task that demands input from an expert or maintainer.

The refactoring of the second function pair, however, was successful. We submitted this refactoring as a patch to the AMD 
Display driver to obtain feedback from the maintainers. The process of sending the patch took more time than expected.
We initially submitted the patch on August 9th, 2024, but we did not receive replies. As a result, we had to resend 
it on October 9th, 2024 and received an initial response on November 3rd, 2024.

We understood new details about the driver evolution processes in discussions with our partner from the AMD Display driver.

Regarding the patch content, only minor changes were requested to align with code 
style and best practices that were initially unknown to us. More details of the 
patch can be found in Appendix \ref{app:feedback}.

>>>>

The initial plan of this research was to approach and sent a large number of mitigations to
collect a broad number of artifacts from the driver community, but the slow process to send a patch
makes this plan unviable. Thus, we opted to redirect this research to work with students and refactoring duplicated function
pairs found by the tool directly in the IIO and the AMD Display driver, without passing through the whole process 
of the systematic approach proposed.
This alternative allowed us to collect artifacts from the students experiences and findings, and not be 
to dependable of the iteraction with the kernel community.
A additional advantage is to accelerate the iteraction with the kernel community
because we are not applying the systematic approach, which translate to smaller code changes.

### 2025-05-25 conteudo/results/students.tex
add some work from 24-05

Appendix (CREATE APPENDIX AND ADD HERE X) has references
to the artifacts discussed in this section, including the duplication function pairs that students 
mitigate, the students's reports, the patch students sent to the linux kernel, and the 
answers of the questionary asked to the student that worked in the AMD Display driver and the 
teaching assistant that used the tool to find duplications in the IIO.

@@

Table (CREATE TABLE AND ADD HERE X) summarizes the mitigation approachs of the students in terms 
of lines changed and refactor methods used. BLABLABLA

>>>>

Appendix \ref{app:stuart} has references
to the artifacts discussed in this section, including the duplication function pairs that students 
mitigate, the students's reports, the patch students sent to the linux kernel, and the 
answers of the questionary asked to the student that worked in the AMD Display driver and the 
teaching assistant that used the tool to find duplications in the IIO.

@@

Table \ref{tab:stu} summarizes the mitigation approachs of students's groups in terms 
of lines changed and refactor methods used. A total of seven groups applied the parameterize refactor method,
and two groups applied the extractor method on their approaches to mitigate the duplications. 
The other three groups does not applied any refactor methods presented in the literature. 
We discuss the groups's approaches and the subsystem maintainers opinions on the patches that students
sent.

Groups one and four have worked in similar duplications, where there are two functions returning a boolean mirroring
the results. That means, when the first function returns true, the second function returns false, and when the first
functions returns false, the second function return false. The group one have a simple approach and the maintainers 
approved the changes quickly. The group four attempted to mitigate the duplication with macros, which is a method 
found in the code file they were working. The maintainers disagree with the proposed approach, arguing 
that the proposed changes decreases the code readability and asked for a approach similar to the group one. 
The group four submitted the patch after coding the approach proposed by the maintainers.

Group three, six and ten created a struct to encapsulate the parameters to mitigate the duplication when 
applying the parameterize method. The group three mitigation was send to the maintainers and the approved 
without significant comments, while the patch from group ten was reject by the same maintainer that approved
the patch from group three. The maintainer justification to reject the patch is that group ten refactored 
code files that has simple context, threfore adding a layer of abstraction to remove a duplication 
deacrease the code readability without clear gains on the codebase quality.

Group seven and eight used the extract method to approach the mitigations. The group eight send the 
mitigations to the maintainers and the code changes was approved with minor requested changes
about code style and smaller errors fix. For the group seven mitigation, the maintainers reject 
their patch because the code changes adds layers of abstractions on the code, turning it more complex
and harder to read. The maintainers proposed to the group seven to remove the duplicated functions 
and insert the duplicated code multiple times were the functions were being called. 
This suggestion we could think that it is a bad practice for code quality, as adds more complexity and 
responsibality to a bigger function, it is view as the best choice for removing layers of abstraction 
and making the code easier to read.

Group nine and eleven faced similar issue as us, where configurations files
designed as macros complicates to approach the mitigations. Group nine applied the parameterize method
to not need to know the configurations in the generic function, which is reasonable as the duplication
approach only depends on one configuration value. Group eleven worked with multiple configuration values,
and to refactor the duplication they proposed to create a bigger macro that encapsulates the other macros,
which is a clever solution but adds abstraction and complexity on macros, which makes the code harder
to debug and read. We do not get the feedback of the maintainers on the group eleven approach to know their 
opinions.

Group twelve is the student that executed the tool and pass throught the process of findings duplications.
The student found a duplication in the driver that happens accross eighteen code files, creating a patch
of removing four hundreds lines of duplicated code. 
The student send the mitigations to the maintainers and their feedback pointed to fix warning in the code, 
without arguing about removing the duplications.

Analysing the students approaches for the duplications and the maintainers feedback, 
we saw that people without previous experience in the linux kernel development could approach 
duplications found by the ArKanjo tool and become contributors to the kernel. We observed that 
not all duplications are view as code with bad quality, with the maintainers analysing the trade off
between what is the purpose of the duplicated code, the code readability, and the levels of abstraction 
added to address the duplications.

### 2025-05-25 conteudo/results/table_students.tex
add some work from 24-05

\textbf{Mitigation} & \textbf{Subsystem} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
\\ \hline

1 & IIO & 0 & 0 & None \\ \hline
2 & IIO & 0 & 0 & None \\ \hline
3 & IIO & 0 & 0 & None \\ \hline
4 & IIO & 0 & 0 & None \\ \hline
5 & IIO & 0 & 0 & None \\ \hline
6 & IIO & 0 & 0 & None \\ \hline
7 & IIO & 0 & 0 & None \\ \hline
8 & IIO & 0 & 0 & None \\ \hline
9 & IIO & 0 & 0 & None \\ \hline
10 & IIO & 0 & 0 & None \\ \hline
11 & IIO & 0 & 0 & None \\ \hline
12 & AMD & 0 & 0 & None \\ \hline

>>>>

\textbf{Group} & \textbf{Subsystem} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
\\ \hline

1 & IIO & 7 & 14 & None \\ \hline
2 & IIO & 23 & 24 & Parameterize method \\ \hline
3 & IIO & 12 & 30 & Parameterize method \\ \hline
4 & IIO & 2 & 7 & None \\ \hline
5 & IIO & 17 & 70 & Parameterize method \\ \hline
6 & IIO & - & - & Parameterize method \\ \hline
7 & IIO & 21 & 22  & None \\ \hline
8 & IIO & - & - & None \\ \hline
9 & IIO & 33 & 52 & Parameterize method \\ \hline
10 & IIO & 32 & 58 & Parameterize method \\ \hline
11 & IIO & - & - & Parameterize method \\ \hline
12 & AMD & 89 & 489 & None \\ \hline
